Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-10-29 Cost_manuscript.Rmd
draft of intro number 1

Outline:
- Effort and delay tend to be aversive, and neuroeconhas traditionally examines them as such
- But there is a flipside of effort: it can be attractive
- One issue is that even for studies that compare across demands, they do so by quantifying discounting curves
- A less explored avenue that might get at this though is the comparison among demands
- Ecological foraging paradigms can help us with this approach, allowing for people to search for their preferred demands
- In addition, foraging rests on a rich modeling history
- We can model whether preferences for demands are due to either 1) modulating cost or embuing value relative to optimal behavior; 2) whether people attend to focal decision elements or global environmental attributes when they evaluate demanding offers
- In this study....

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. The intuitive reaction to this evaluation is to find demands aversive. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness; Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what intrinsic features of an effortful task can make it attractive beyond its associated outcome, and how contextual factors mold their perceived costs (see Inzlicht, Shenhav, and Olivola, 2018, for a review on the attractiveness of effort).

  What is often missing from these accounts is the common case of choosing among behavioral demands, partly due to the scope of traditional experimental designs. Studies tend to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. However, flexible preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available (Olivola and Shafir, 2013).

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019). Foraging has attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [49]. The paradigm requires the decision maker to choose whether to engage with a single presented offer or skip it in search of a better one. A key advantage of foraging paradigms is that people can rapidly adopt stable, rewardmaximizing strategies [50]. Add that Fawcett bit and modeling…

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions. First, the existence of a well-defined optimal choice strategy in foraging designs was useful for identifying when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands over two experiments let us probe whether their cost is intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences are domain general or task-specific, and in the case of the latter, isolate factors that influence preferences beyond the often cited associative relationship between effort and outcome (e.g. features of an effortful task that makes it more attractive).

  Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups underaccepted, partially matching our predictions and suggesting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the perceived time of trials, a phenomenon that could be recalibrated as individuals experienced multiple forms of demands (a detailed description of our preregistered hypotheses can be found on https://osf.io/2rsgm/). Our results expand our understanding of the attractiveness of effort, highlighting a role for features in the decision environment that are independent from a pure association between effort and reward.

####-------######
  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).
  In this study, we will test the hypothesis that preferences will vary depending on whether individuals face a single or multiple types of behavioral demands, and identify what factors drive these choices. Based on two foraging experiments and computational modeling, we will show that 1) preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.
  In this study, we will behaviorally and computationally test the hypothesis that the attractiveness of a reward that requires effort or delay depends on whether rewards associated with alternative demands are available in the same context. Using a foraging paradigm will enable us to 1) assess whether people integrate the statistics of the environment to generate strategies that maximize their earnings; 2) compare how the presence of effort or delay shifts such strategies, as reflected by acceptance rates; 3) examine if acceptance rates are further altered by foraging for costs in addition to rewards; and 4) test a combination of foraging (Constantino & Daw, 2015; Stephens & Krebs, 2019; Fawcett, McNamara, & Houston, 2012) to examine whether preferences for demands reflect a subjective reduction of cost or boost of value, and determine any factors besides contextual manipulations that participants consider in order to adjust their choices.

studies have not focused on how effort and delay might interact in the common case of choosing among demands.

  Grant: Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. Demands for both time and effort have been identified as costs that detract from the net subjective value of rewards. Decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [1–9]. Excessive aversion to delay and effort can manifest as impulsivity and apathy, respectively [1,10–12], fluctuations in which are hallmark symptoms of various mental illnesses and substance use disorders [13]. However, we do not yet fully understand 1) how the brain transforms these two parallel forms of discounting into seemingly opposite behaviors; 2) whether the same mechanism applies to different types of effort (e.g. physical versus cognitive); or 3) the ways in which these distinct forms of cost interact in individuals. In this project, I plan to combine individualized brain mapping with computational modeling to examine how an individual’s functional brain network organization relates to the subjective costs of physical effort, cognitive effort, and pure delay.
  Behavioral distinctions between effort and delay costs: The process of engaging with a costly reward can be divided into three stages: evaluation (when costs and benefits of a prospect are integrated), pursuit (i.e. waiting or exerting effort), and receipt of the reward. An ever-increasing literature in psychology, neuroscience, and economics has focused on the first stage to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated demand [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. Investigations of the pursuit and receipt stages have been comparatively limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards [3,21,22].
  The evidence reviewed above has two main limitations. First, studies have not focused on how effort and delay might interact in the common case of choosing among demands. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. 
  It is unknown how effort demands discount value in isolation versus when alternative demands are available. The friction that we perceive from these demands could depend on what we could be performing instead [61].

IDEAS:

   Studies have often focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. (other frame: What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands). Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

So going back to our question of the cost and value of effort, maybe the distinction has something to do with the ways in which we can allocate time and effort (work on this, including how foraging can provide a reference point to evaluate the value-modulatory effects of demands, and maybe put it before foraging is explained?).

  Modeling: mention that the rich history of foraging modeling can help us pinpoint a number of things, like the putative cost or attractiveness of a demand, whether preferences emerge from focal considerations of the value of a trial or a global estimate of what alternative way could be used to obtain rewards. Foraging and its modeling can provide a common framework to examine differences, rather than relying on matching disc functions.

  Clinical: Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- I guess you can mention that the wait group acts as a baseline of foraging behavior.
- Re-read Mikey's effort value paper to see if there is anything there that can allow us to transition into foraging. Like, are there elements that make cog effort valuable that are well encapsuled by foraging paradigms?
- Pending questions: explain logistic coefficients in exponentiated form? (i.e. X% more/less likely than...)

>>>>

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002; Mochon, Norton, and Ariely, 2012; Hernandez Lallement et al., 2014), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness, Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what intrinsic features of an effortful task can make it attractive outside of its associated outcome, and how contextual factors mold their perceived costs (see Inzlicht, Shenhav, and Olivola, 2018, for a review on the attractiveness of effort). Understanding these additional qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available (Olivola and Shafir, 2013). Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time (OC; Fawcett et al., 2012; Kurzban et al, 2013; Otto and Daw, 2019). This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort (Inzlicht, Shenhav, and Olivola, 2018).

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Stephens & Krebs, 2019). Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors (Hayden 2018; Mobbs, Trimmer, Blumstein, & Dayan, 2018), and offers a number of desirable qualities. First, it allows for the estimation of reward maximizing choice patterns based on the opportunity cost of time (Charnov, 1975; Kacelnik and Krebs, 1977), which people can readily approximate and persistently follow (Constantino and Daw, 2015 ADD). One can then interpret observed deviations in acceptance rates from such optimal behavior in addition to comparing them among demands. Second, the sequential nature of foraging allows individuals to emphasize different elements of the decision context. For example, the cost (or value) of effort or delay could globally depend on the availability of alternative demands (as suggested above), impacting the perceived opportunity cost of time. Alternatively, it could be intrinsic and robust to such manipulation, and/or depend on task-specific factors that adapt once individuals experience different forms of behavioral costs. Foraging rests on an abundant modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions. First, the existence of a well-defined optimal choice strategy in foraging designs helped us identify when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands over two experiments allowed us to asses whether their cost was intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences were domain general or task-specific, and in the case of the latter, isolate factors that influence preferences beyond the often cited associative relationship between effort and outcome (e.g. features of an effortful task that makes it more attractive).

  Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment, or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups underaccepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands (a detailed description of our preregistered hypotheses can be found on https://osf.io/2rsgm/). Our results expand our understanding of the attractiveness of effort, highlighting a role for features in the decision environment that are independent from a pure association between effort and reward, and point to characteristics that can be leveraged in order to motivate effort engagement in diverse everyday scenarios.

####-------######

  Clinical: Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
